---
title: "Advanced: Change MAgPIE GAMS Code"
subtitle: MAgPIE model development team (magpie@pik-potsdam.de)
date: "`r format(Sys.time(), '%d %B, %Y')`"
author: Florian Humpenöder (humpenoeder@pik-potsdam.de)
output:
  rmarkdown::github_document:
    toc: yes
  html_notebook:
    code_folding: none
---

### 1 Introduction
MAgPIE has a modular concept. Each module (e.g. `pasture`) can have several realizations (e.g. `dynamic` and `static`). The purpose of these realization is a) to maintain the current default model behavior and b) to keep the model operational while developing a new realization. A typical use case is the extension of a realization by a specific feature. In this case, one would copy the current default realization, rename it properly, and apply the wanted changes. Then, the model behavior can be compared between the two realizations. Eventually, the new realization might become the new default at some point and the old realization is deleted.

### 2 Learning objectives
This tutorial shows how to add a new realization to a module in the MAgPIE model. To illustrate the different steps, we will expand the `urban` land module. In the current MAgPIE master, the `urban` land module has only a `static` realization. In this tutorial, we will add a `dynamic` realization, which changes urban land based on population growth. 

### 3 Adding a new realization
We want to add a new realization to the `urban` land module. The `urban` land module is located here: `modules/34_urban`.

#### Add a new realization by duplicating an existing one
Duplicate the `static` folder and rename it to `dynamic`. Now we need to add and edit files in the `dynamic` folder. Add the following files by duplicating the `presolve.gms` file: `declarations.gms`, `equations.gms` and `preloop.gms`. Open each new file and delete the copied code from `presolve.gms`, but keep the copyright header. Now add the following content to these files, and modify `presolve.gms` and `realization.gms`.

##### declarations.gms
```{r, eval=FALSE}
equations
 q34_urban(j)       		              urban land (mio. ha)
;

parameters
 p34_pop_growth(t,i) population growth factor between time steps (1)
;
```

##### equations.gms
```{r, eval=FALSE}
q34_urban(j2)..
 vm_land(j2,"urban") =e= 
 pcm_land(j2,"urban") * sum((ct,cell(i2,j2)), p34_pop_growth(ct,i2));
```

##### preloop.gms
```{r, eval=FALSE}
loop(t$(ord(t) > 1),
 p34_pop_growth(t,i) = im_pop(t,i)/im_pop(t-1,i);
);
```

##### presolve.gms
```{r, eval=FALSE}
*vm_land.fx(j,"urban") = pcm_land(j,"urban");
vm_carbon_stock.fx(j,"urban",c_pools) = 0;
```

##### realization.gms
```{r, eval=FALSE}
*' @description In the dynamic realization, urban land expands based on population growth.
*' Carbon stocks are fixed to zero because
*' information on urban land carbon density is missing.

*' @limitations Carbon stocks are assumed zero.
```

#### Update the code
To include the new realization `dynamic` properly into the GAMS code we run a specific R command in the main folder. First navigate in your command line to the MAgPIE main folder, then open a new R session (type `R` followed by ENTER), and then copy-paste the following R commands:
```{r, eval=FALSE}
gms::update_fulldataOutput()
gms::update_modules_embedding()
```
These two commands will add a `postsolve.gms` file, and update the `realization.gms` and `declarations.gms` files.
Hint: If you change, add or delete variables/parameters always run these commands to avoid GAMS compilation errors.

Run codeCheck to check if all module interfaces exist. 
```{r, eval=FALSE}
gms::codeCheck()
```

Now you can quit the R session with `q()`.

### 4 Testing a new realization
#### Start a model run
For a quick GAMS test, we simply set the new realization in the file `main.gms` in line 256 (replace `static` by `dynamic`). We can then check if the model compiles correctly with this command evoked from the command line.
```{r, eval=FALSE}
gams main.gms action=C
```
If you get compilation errors, you have to resolve these first. Look into the `main.lst` file. It will tell you what kind of error occurred. 

To make our test run as fast as possible, we reduce the number of time steps to 3 and deactivate elastic food demand. 
For this, we change `$setglobal c_timesteps  coup2100` in main.gms in line 224 to `$setglobal c_timesteps  quicktest`, and change the value of `s15_elastic_demand` from 1 to 0 in line 74 in `modules/15_food/input.gms`.

Now we can start a test run with this command. This can take a while depending on the resources of your machine.
```{r, eval=FALSE}
gams main.gms
```
GAMS will create a `fulldata.gdx` file in the main folder.

For starting a productive model run, we would have to change the config file `config/default.cfg` in line 489 (replace `static` by `dynamic`). We could now start a model run with `Rscript start.R -> 1: default -> 1: Direct execution`.

#### Check the results
Start a new R session in the MAgPIE main folder, and execute these commands.
```{r, eval=FALSE}
options(digits=2)
library(magpie4)
gdx <- "fulldata.gdx"
land(gdx,level="glo",type="urban")
land(gdx,level="reg",type="urban")
```

